### 2017-11-01 submission/manuscript.Rmd
update result section name and add methods on CRC associated Genera OTU model
add results on whole community versus specific OTUs

***Using the Whole Community Increases AUC:***

@@

***Study Analysis Overview:***
Alpha diversity was first assessed for differences between controls and adenoma versus cancer and controls versus adenoma. We analyzed the data using linear mixed-effect models, and relative risk. Beta-diversity was then assessed for each inidividual study. Next, four specific CRC-associated genera (*Fusobacterium*, *Parvimonas*, *Peptostreptococcus*, and *Porphyromonas*) were assessed for differences in relative risk. We then built Random Forest models based on all genera or the select CRC-associated genera. The models were trained on one study then tested on the remaining studies for every study. The data was split between feces and tissue samples. Within the tissue groups the data was further divided between matched and unmatched tissue samples. Both prediction for adenoma and carcinoma were tested. This same approach was then applied at the OTU level with the exception that instead of testing on the other studies a 10-fold cross validation was utilized and 100 different models were created based on random 80/20 splitting of the data to generate a range of expected AUCs. The power of each study was assessed for and effect size ranging from 1% to 30%. An estimated sample n for these effect sizes was also generated based on 80% power.

>>>>

***Using the Whole Community Increases Model AUC:***
For both fecal and tissue samples (matched and unmatched) there was a decrease in AUC when only OTUs from the CRC associated genera are used [Figure 4 & 5]. This decrease is observed in both adenoma and carcinoma groups [Figure 4 & 5]. The genus models generally had similar trends as observed for the OTU based models with the full genera models performing better then the CRC associated genera models [Figure S4-S4]. Both genus models perform similarily in their ability to be able to predict lesion (adenoma or carcinoma) with carcinoma having a higher AUC then adenoma [Figure S6-S8]. Matched tissue samples for those with carcinoma had an AUC that was more similar to the adenoma models [Figure S6A, S7B, & S8] then carcinoma models [Figure S6B & S7A].

@@

***Study Analysis Overview:***
Alpha diversity was first assessed for differences between controls and adenoma versus cancer and controls versus adenoma. We analyzed the data using linear mixed-effect models, and relative risk. Beta-diversity was then assessed for each inidividual study. Next, four specific CRC-associated genera (*Fusobacterium*, *Parvimonas*, *Peptostreptococcus*, and *Porphyromonas*) were assessed for differences in relative risk. We then built Random Forest models based on all genera or the select CRC-associated genera. The models were trained on one study then tested on the remaining studies for every study. The data was split between feces and tissue samples. Within the tissue groups the data was further divided between matched and unmatched tissue samples. Both prediction for adenoma and carcinoma were tested. This same approach was then applied at the OTU level with the exception that instead of testing on the other studies a 10-fold cross validation was utilized and 100 different models were created based on random 80/20 splitting of the data to generate a range of expected AUCs. For OTU based models the CRC Associated Genera included all OTUs that had a taxonomic classification to *Fusobacterium*, *Parvimonas*, *Peptostreptococcus*, or *Porphyromonas*. The power of each study was assessed for and effect size ranging from 1% to 30%. An estimated sample n for these effect sizes was also generated based on 80% power.

### 2017-11-01 submission/manuscript.tex
add results on whole community versus specific OTUs

\textbf{\emph{Using the Whole Community Increases Model AUC:}}

>>>>

\textbf{\emph{Using the Whole Community Increases Model AUC:}} For both
fecal and tissue samples (matched and unmatched) there was a decrease in
AUC when only OTUs from the CRC associated genera are used {[}Figure 4
\& 5{]}. This decrease is observed in both adenoma and carcinoma groups
{[}Figure 4 \& 5{]}. The genus models generally had similar trends as
observed for the OTU based models with the full genera models performing
better then the CRC associated genera models {[}Figure S4-S4{]}. Both
genus models perform similarily in their ability to be able to predict
lesion (adenoma or carcinoma) with carcinoma having a higher AUC then
adenoma {[}Figure S6-S8{]}. Matched tissue samples for those with
carcinoma had an AUC that was more similar to the adenoma models
{[}Figure S6A, S7B, \& S8{]} then carcinoma models {[}Figure S6B \&
S7A{]}.
